---
title: "MMRM Analysis using the `tern.mmrm` Package"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction

This document provides an overview of fitting a **Mixed Model for Repeated Measures (MMRM)** using the `tern.mmrm` R package. MMRM is a specific linear mixed effects model that includes fixed effects for the following variables:
- **Treatment Arm**
- **Categorical Visit**
- **Treatment by Visit Interaction**
- Other covariates for adjustment (e.g., age, gender)

The model uses maximum likelihood estimation and is particularly useful for analyzing repeated measures data.

---

## Setup

Load the necessary libraries and data.

```{r}

# Load additional packages
library(dplyr)
library(ggplot2)
if (!require("lme4")) {
  install.packages("lme4")
}
library(lme4)
library(dplyr)

```


```{r}
data %>% select (-FatherMyop, - MotherMyop)

```

```{r}
# Fit the LMM with random intercepts for subjects
mmrm_model <- lmer(
  SER ~ TrtGroup * Visit + MotherMyop + FatherMyop + Sex + Race + EyeColor +
    AgeAsofEnrollDt + genetic + (1 | SubjectID),
  data = data,
  REML = TRUE
)
# Summarize the model results
summary(mmrm_model)
```


