---
title: "MMRM Analysis using the `tern.mmrm` Package"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction

This document provides an overview of fitting a **Mixed Model for Repeated Measures (MMRM)** using the `tern.mmrm` R package. MMRM is a specific linear mixed effects model that includes fixed effects for the following variables:
- **Treatment Arm**
- **Categorical Visit**
- **Treatment by Visit Interaction**
- Other covariates for adjustment (e.g., age, gender)

The model uses maximum likelihood estimation and is particularly useful for analyzing repeated measures data.

---

## Setup

Load the necessary libraries and data.

```{r}

# Load additional packages
library(dplyr)
library(ggplot2)
if (!require("lme4")) {
  install.packages("lme4")
}
library(lme4)
library(dplyr)

```

```{r}
load("data.RData")

clean_data <- data %>%
  # First reclassify Race
  mutate(
    Race = case_when(
      Race == "Black/African American" ~ "Black",
      Race == "White" ~ "White",
      TRUE ~ "Other"  # All other categories combined
    ),
    # Convert Race to factor with specific reference level
    Race = factor(Race, levels = c("Black", "White", "Other")),
    # Then clean Visit
    Visit = case_when(
      Visit == "Enrollment" ~ "Month0",
      Visit == "Run-in FU Randomization" ~ "Randomization",
      Visit == "Month 6 Visit" ~ "Month6",
      Visit == "Month 12 Visit" ~ "Month12", 
      Visit == "Month 18 Visit" ~ "Month18",
      Visit == "Month 24 Visit" ~ "Month24",
      Visit == "Month 30 Visit" ~ "Month30"
    ),
    # Divide AL into four subgroups using specific breakpoints
    AL_group = cut(
      AL,
      breaks = c(-Inf, 24.5, 26, 27.5, Inf),  # Specify breakpoints
      labels = c("Group1", "Group2", "Group3", "Group4"),  # Subgroup labels
      include.lowest = TRUE
    )
  ) %>%
  # Create binary indicator variables for each subgroup
  mutate(
    AL_group1 = as.integer(AL_group == "Group1"),
    AL_group2 = as.integer(AL_group == "Group2"),
    AL_group3 = as.integer(AL_group == "Group3"),
    AL_group4 = as.integer(AL_group == "Group4")
  ) %>%
  filter(Visit != "Randomization") %>% 
  mutate(
      Visit_numeric = case_when(
        Visit == "Month0" ~ 0,
        Visit == "Month6" ~ 6,
        Visit == "Month12" ~ 12,
        Visit == "Month18" ~ 18,
        Visit == "Month24" ~ 24,
        Visit == "Month30" ~ 30
      )
    )

# Split into four separate datasets based on subgroups
group1_data <- clean_data %>% filter(AL_group1 == 1) %>% select(-AL_group1, -AL_group2, -AL_group3, -AL_group4)
group2_data <- clean_data %>% filter(AL_group2 == 1) %>% select(-AL_group1, -AL_group2, -AL_group3, -AL_group4)
group3_data <- clean_data %>% filter(AL_group3 == 1) %>% select(-AL_group1, -AL_group2, -AL_group3, -AL_group4)
# group4_data <- clean_data %>% filter(AL_group4 == 1) %>% select(-AL_group1, -AL_group2, -AL_group3, -AL_group4)



```

```{r}
# Fit the LMM with random intercepts for subjects
library(nlme)
lmm_model <- lmer(
  AL ~ TrtGroup * as.factor(Visit_numeric) + Sex + Race + EyeColor +
    AgeAsofEnrollDt + genetic + (1 | PtID),
  data = group3_data,
  REML = TRUE
)


        
# Summarize the model results
summary(lmm_model)

library(broom.mixed)  # To tidy model summaries
library(dplyr)

# Function to summarize fixed effects and calculate p-values
summarize_lmer_results <- function(model) {
  # Extract fixed effects and their statistics
  summary_df <- broom.mixed::tidy(model, effects = "fixed") %>%
    mutate(
      p_value = 2 * pt(-abs(statistic), df = df.residual(model)),  # Calculate p-value
      significance = case_when(
        p_value < 0.001 ~ "***",
        p_value < 0.01 ~ "**",
        p_value < 0.05 ~ "*",
        TRUE ~ "ns"
      )
    )
  
  # Return the summarized table
  summary_df %>%
    select(term, estimate, std.error, statistic, p_value, significance) %>%
    arrange(p_value)
}

# Example usage
model_summary <- summarize_lmer_results(lmm_model)
print(model_summary)

```


```{r}

# Fit the LMM with random intercepts for subjects
library(nlme)
lmm_model1 <- lmer(
  AL ~ TrtGroup*as.factor(genetic)*Visit_numeric + Sex + Race + EyeColor +
    AgeAsofEnrollDt 
  + ( Visit_numeric | PtID),
  data = group1_data,
  REML = TRUE
)
summarize_lmer_results(lmm_model1)


lmm_model2 <- lmer(
  AL ~ TrtGroup*as.factor(genetic)*Visit_numeric + Sex + Race + EyeColor +
    AgeAsofEnrollDt 
  + ( Visit_numeric | PtID),
  data = group2_data,
  REML = TRUE
)
summarize_lmer_results(lmm_model2)

lmm_model3 <- lmer(
  AL ~ TrtGroup*as.factor(genetic)*Visit_numeric + Sex + Race + EyeColor +
    AgeAsofEnrollDt 
  + ( Visit_numeric | PtID),
  data = group3_data,
  REML = TRUE
)
summarize_lmer_results(lmm_model3)

```
```{r}

# Load necessary libraries
library(lme4)
library(broom.mixed)
library(dplyr)

# Define a function to fit and summarize LMM for a given dataset
fit_and_summarize_lmm <- function(data, group_name) {
  # Fit the model
  lmm_model <- lmer(
    AL ~ TrtGroup * as.factor(genetic) * Visit_numeric + Sex + Race + EyeColor +
      AgeAsofEnrollDt + (Visit_numeric | PtID),
    data = data,
    REML = TRUE
  )
  
  # Summarize the model
  summary_df <- broom.mixed::tidy(lmm_model, effects = "fixed") %>%
    mutate(
      p_value = 2 * pt(-abs(statistic), df = df.residual(lmm_model)),
      significance = case_when(
        p_value < 0.001 ~ "***",
        p_value < 0.01 ~ "**",
        p_value < 0.05 ~ "*",
        TRUE ~ "ns"
      )
    )
  
  # Add group name for clarity
  summary_df <- summary_df %>%
    mutate(group = group_name)
  
  return(summary_df)
}

# Fit and summarize models for all groups
results_group1 <- fit_and_summarize_lmm(group1_data, "Group 1")
results_group2 <- fit_and_summarize_lmm(group2_data, "Group 2")
results_group3 <- fit_and_summarize_lmm(group3_data, "Group 3")

# Combine results into a single table
all_results <- bind_rows(results_group1, results_group2, results_group3)

# Print the summarized results
print(all_results)
```
